\section{Algorithm and Analysis}\label{analysis}

\subsection{Algorithms for MFEC-MAR (Fairness in WSNs)}
In this section we will explain our two step approach to solving the fairness problem in WSNs.

\subsubsection{Connecting Nodes}
This section explains how to connect and initially disconnected set of nodes using the Steiner Minimum Tree (SMT) algorithm.
\subsubsection{Optimizing Node Location}
In this section we explain an extension to the SMT problem to include a fairness measure across the network of nodes.
The explanation will include the moving of Steiner Points to optimal locations as well as the introduction of additional resources (nodes) to achieve optimality. The discussion includes an overview of the trade-off between additional resources and  fairness and illustrates results of the optimization operations across multiple dimensions (such as minimizing the deployment of additional resources, maximizing fairness while minimizing power consumption).





The section compares our approach the 3-approximation algorithm for minimal power consumption of WSNs.
\subsection{Theoretical Analysis}

Our approach will be compared to the 3-approximations time and space usages from a theoretical standpoint.

\subsection{Experimental Analysis}
The theoretical analysis will be verified by an experimental analysis.

\linesnumbered
\begin{algorithm}
\dontprintsemicolon
\SetKwData{FunctionSignature}{MakeFair}
\FunctionSignature{T,S}
\\
\KwIn{$G=(X,U)$ such that $G^{tc}$ is an order.}
\KwOut{$G’=(X,V)$ with $V\subseteq U$ such that $G’^{tc}$ is an
interval order.}
\Begin{
$V \longleftarrow U$\;
$S \longleftarrow \emptyset$\;
\For{$x\in X$}{
$NbSuccInS(x) \longleftarrow 0$\;
$NbPredInMin(x) \longleftarrow 0$\;
$NbPredNotInMin(x) \longleftarrow |ImPred(x)|$\;
}
\For{$x \in X$}{
\If{$NbPredInMin(x) = 0$ {\bf and} $NbPredNotInMin(x) = 0$}{
$AppendToMin(x)$}
}
\lnl{InRes1}\While{$S \neq \emptyset$}{

remove $x$ from the list of $T$ of maximal index\;
\lnl{InRes2}\While{$|S \cap ImSucc(x)| \neq |S|$}{
\For{$ y \in S-ImSucc(x)$}{
\{ remove from $V$ all the arcs $zy$ : \}\;
\For{$z \in ImPred(y) \cap Min$}{
remove the arc $zy$ from $V$\;
$NbSuccInS(z) \longleftarrow NbSuccInS(z) - 1$\;
move $z$ in $T$ to the list preceding its present list\;
\{i.e. If $z \in T[k]$, move $z$ from $T[k]$ to
$T[k-1]$\}\;
}
$NbPredInMin(y) \longleftarrow 0$\;
$NbPredNotInMin(y) \longleftarrow 0$\;
$S \longleftarrow S - \{y\}$\;
$AppendToMin(y)$\;
}
}
$RemoveFromMin(x)$\;
}
}
\caption{IntervalRestriction\label{IR}}
\end{algorithm}
