
\documentclass[11pt,conference]{IEEEtran} 
%\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage[tight,footnotesize]{subfigure}
\usepackage{url}
\usepackage{algorithm2e}
\usepackage{verbatim}

\begin{document}

% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Fair Energy Consumption in Wireless Sensor Networks}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Niels Kasch*}
\IEEEauthorblockA{Email: nkasch1@umbc.edu}
\and
\IEEEauthorblockN{Dave Feltenberger*}
\IEEEauthorblockA{Email: dfelten1@umbc.edu\\\\
*Department of Computer Science and\\Electrical Engineering\\
University of Maryland, Baltimore County\\
Baltimore, MD 21250\\
April 27$^{\textrm{th}}$, 2009\\
}
\and
\IEEEauthorblockN{Fatih Senel*}
\IEEEauthorblockA{Email: fsenel1@umbc.edu}
}

\maketitle



\begin{abstract}
We introduce the notion of \textit{fairness} in power consumption of wireless sensor networks (WSNs). 
Herein, \textit{fairness} in power consumption refers to an approximately equal power consumption rate among all nodes in a WSN. This topic has been largely overlooked by the literature, but has a significant impact in the design of WSNs with fixed sensor nodes and resource scarce flexible relay nodes.

We propose an approximation algorithm with the aim to introduce fairness to power consumption in WSNs. While overall power consumption has been studied previously, our algorithm considers fairness of power consumption rates a priority in WSN design. Our algorithm introduces additional, limited resources (i.e. relay nodes) to a fixed network. These resources may be moved to achieve minimal overall power consumption and maximum fairness. Hence, we give a multi-objective optimization algorithm for (1) maximum fairness in power consumption, (2) minimal additional resources and (3) minimal power consumption of WSNs.

Experimental analysis of randomly generated WSNs show that the introduction of a small numbers of relay nodes (as compared to sensor nodes) results in a significant increase in fairness of power consumption. However, as further relay nodes are added, we encounter a diminishing return in the increase of fairness. Hence, we pose the question of optimality considering the trade off between fairness and relay nodes and give a discussion for the experimental determination of such optimality.

We briefly illustrate our WSN fairness simulation software that utilizes the algorithm and visualizes the process of optimizing a WSN for fairness.
\end{abstract}

\begin{keywords}
Wireless Sensor Network, Energy Conservation, Fairness, Energy Consumption
\end{keywords}

\IEEEpeerreviewmaketitle


\input{introduction}
\input{relatedWork}
\input{approach}
\input{analysis}
\input{results}
\input{software}


\section{Conclusion}\label{conclusion}

Approximating fairness in sensor networks has had little to no attention in the literature.  This is perhaps because in many cases, only total power consumption matters.  When dealing with sensor networks where changing, repairing, or working on sensors is invasive-for instance a sensor network on a person, or a network in a difficult to reach destination-fairness becomes important.  As discussed throughout the paper, the reason fairness is important is to ensure node failures occur approximately at the same time.  With approximate fairness, this becomes possible with regard to battery failure, as nodes will deplete power reserves at roughly the same rate.  With node failures occurring at the same time, it allows a network to operate under better-understood and more predictable conditions, as well as helping to avoid two undesirable properties.  First, invasive node replacement procedures do not occur as frequently as if single nodes were allowed to consume power at significantly different rates: with unfair conditions, certain nodes may fail significantly sooner than others, thereby requiring replacement procedures more frequently.  Second, because the procedure is invasive, it is desirable to replace all nodes during the same procedure; under this model, fairness means less waste, since there will be fewer nodes with significant battery power left, while unfair networks potentially waste nodes with significant battery power if all nodes are replaced at the same time.

The fairness approximation procedure discussed above is difficult to measure in any absolute sense.  The reason is twofold: first, there are no baseline comparisons to be made, as the authors could not find fairness studies in the literature; second, and more importantly, the context of the application changes characteristics significantly.  Larger surface areas require more relay nodes to achieve fairness.  Similarly, when the number of terminal nodes is low, more relay nodes are needed to achieve fairness.  Conversely, smaller surface area or more terminal nodes tend to require fewer relay nodes.  These observations are both intuitive and experimentally shown.  This is a significant advantage of the fairness procedure described in this paper: consistent results that follow intuition.  To evaluate the number of relay nodes necessary, it is advisable, therefore, to run simulations and find a $k$ value that best meets the requirements of the application.

\section{Future Work}\label{futureWork}

There are a number of areas touched on in this paper that could be studied more extensively in future work.  The most important consideration that was ignored, which in practical applications will almost definitely be an important measure, is reducing overall power consumption.  The two-fold optimization of optimizing fairness and power consumption would prove very useful in the use cases outlined in this paper, too.

A second future area to consider is connecting all nodes within range and including them in fairness and power measurements.  Currently the algorithm works on a tree without cycles for simplicity, but in any real application of the sensor networks, nodes would need to consider all other nodes within range instead of just one.


\section*{Acknowledgment}

The authors would like to thank Dr. Alan Sherman for his valuable feedback and suggestions related to the design and analysis of the algorithms presented herein and his guidance while writing this manuscript. We are grateful for Neelofer Tamboli for her supporting work regarding this project.



\bibliographystyle{IEEEtran}
\bibliography{IEEEfull,paper}

\onecolumn
\input{appendix}

% that's all folks
\end{document}



